### 2016-09-13 paper/paper.tex
Let's not say this

% vim: set spell spelllang=en tw=100 :

\documentclass[letterpaper]{article}

\usepackage[pass]{geometry}

\usepackage{ijcai16}
\usepackage{times}
\usepackage{complexity}
\usepackage{microtype}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{cleveref}
\usepackage{tikz}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage[ruled,vlined]{algorithm2e}

% \usepackage{showframe}

\makeatletter\@ifpackageloaded{showframe}{
    % show lines down the middle columns too
    % http://tex.stackexchange.com/questions/16199/test-if-a-package-or-package-option-is-loaded
\newlength\Fcolumnseprule\setlength\Fcolumnseprule{0.4pt}\def\@outputdblcol{%
  \if@firstcolumn\global \@firstcolumnfalse \global \setbox\@leftcolumn \box\@outputbox
  \else
    \global \@firstcolumntrue
    \setbox\@outputbox \vbox{\hb@xt@\textwidth{\hb@xt@\columnwidth{\box\@leftcolumn \hss}%
    \vrule \@width\Fcolumnseprule\hfil{\normalcolor\vrule \@width\columnseprule}
    \hfil\vrule \@width\Fcolumnseprule\hb@xt@\columnwidth {\box\@outputbox \hss}}}\@combinedblfloats\@outputpage
    \begingroup\@dblfloatplacement\@startdblcolumn\@whilesw\if@fcolmade \fi{\@outputpage\@startdblcolumn}\endgroup
  \fi
}}{}
\makeatother

\usetikzlibrary{decorations, decorations.pathreplacing, calc, backgrounds,shapes.misc}
%cross
\tikzset{cross/.style={cross out, draw=uofgcobalt, minimum size=2*(#1-\pgflinewidth), inner sep=0pt, outer sep=0pt},cross/.default={1.5mm}}

\definecolor{uofgsandstone}{rgb}{0.321569, 0.278431, 0.231373}
\definecolor{uofglawn}{rgb}{0.517647, 0.741176, 0}
\definecolor{uofgcobalt}{rgb}{0, 0.615686, 0.92549}
\definecolor{uofgpumpkin}{rgb}{1.0, 0.72549, 0.282353}
\definecolor{uofgthistle}{rgb}{0.584314, 0.070588, 0.447059}

\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
\newcommand{\citep}[1]{\cite{#1}}

\theoremstyle{definition}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{example}{Example}

% cref style
\crefname{figure}{Figure}{Figures}
\Crefname{figure}{Figure}{Figures}

\crefname{algocf}{Algorithm}{Algorithms}
\Crefname{algocf}{Algorithm}{Algorithms}

% http://tex.stackexchange.com/questions/22100/the-bar-and-overline-commands
\newcommand{\shortoverline}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}

% Maths operators
\newcommand{\paths}{\operatorname{paths}}
\newcommand{\lessnonind}[1]{\prescript{}{#1}{\rightarrowtail}\ }
\newcommand{\lessind}[1]{\prescript{}{#1}{\hookrightarrow}\ }
\newcommand{\lessmap}[1]{\prescript{}{#1}{\mapsto}\ }
\newcommand{\lesspreceq}[1]{\prescript{}{#1}{\preceq}\ }
\newcommand{\V}{\operatorname{V}}
\newcommand{\N}{\operatorname{N}}
\newcommand{\nds}{\operatorname{S}}
\newcommand{\loopcomp}[1]{\oset[.1ex]{\multimap}{#1}}
% to do comments
\usepackage{color}
\newcommand{\todo}[1]{{\color{red} {?? [}{#1}{]}}}

\makeatletter
\newcommand{\oset}[3][0ex]{%
  \mathrel{\mathop{#3}\limits^{
    \vbox to#1{\kern-2\ex@
    \hbox{$\scriptstyle#2$}\vss}}}}
\makeatother

\title{Between Subgraph Isomorphism and Maximum Common Subgraph}

\author{Anonymous \and Anonymous\thanks{This work was supported by an anonymous funding agency [grant number 12345]} \and Anonymous\thanks{This work was supported by an anonymous funding agency [grant number 67890]} \\
Top Secret Research Lab, Atlantis \\
anonymous@anonymous}

\pdfinfo{
    /Title (Between Subgraph Isomorphism and Maximum Common Subgraph)
    /Author (Anonymous and Anonymous and Anonymous)
}

\begin{document}

\maketitle

\begin{abstract}
    When a small pattern graph does not occur inside a larger target graph, we can ask how to find
    ``as much of the pattern as possible'' inside the target graph. In general, this is known as the maximum
    common subgraph problem, which is much more computationally challenging in practice than
    subgraph isomorphism. We introduce a restricted alternative, where we ask if all but $k$
    vertices from the pattern can be found in the target graph. This allows for the development of
    slightly weakened forms of certain invariants from subgraph isomorphism which are based upon degree and
    number of paths.  We show that when $k$ is small, weakening the invariants still retains much of
    their effectiveness. We are then able to solve this problem on the standard SIP problem
    instances used to benchmark subgraph isomorphism algorithms, despite these instances being
    too large for current maximum common subgraph algorithms to handle. Finally, by iteratively
    increasing $k$, we obtain an algorithm which is also competitive for the maximum common subgraph
    problem.
\end{abstract}

\section{Introduction}
\todo{talk about subgraph isomorphism applications, particularly vision}

The subgraph isomorphism problem is to find a copy of a small \emph{pattern} graph inside a larger
\emph{target} graph. It comes in two forms: in the non-induced variant, edges must be mapped to
edges, but the target may have ``extra edges'', whilst in the induced variant, non-edges may only be
mapped to non-edges. When a pattern cannot be found, we may wish to be given a result which maps as
many vertices of the pattern into the target as possible. In the induced case, this is known as the
maximum common induced subgraph problem (we discuss the non-induced case below). However, although
recent subgraph isomorphism algorithms are comfortable working with graphs with thousands of
vertices, the state of the art for the maximum common subgraph problem becomes computationally
infeasible at only 35 vertices when working with unlabelled graphs
\citep{DBLP:conf/cp/McCreeshNPS16}. This is largely because strong inference, based upon the degrees
of vertices \citep{DBLP:journals/ai/Solnon10} and the distances or paths between them
\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, is possible with subgraph isomorphism,
but not maximum common subgraph, and so the state space in the former is much more restricted,
whilst filtering during search is vastly stronger.

In this work we discuss an intermediate problem, where we must map all but $k$ vertices of the
pattern graph into the target. We show that if $k$ is reasonably small (say, between 1 and 5), then
weakened forms of degree and path based filterings are still effective in pruning the initial search
space and in providing additional constraints respectively. We then show that combining these
techniques leads to a practical algorithm which can scale to work with the families of graphs
commonly used to benchmark subgraph isomorphism algorithms: depending upon the benchmark family, we
can close a substantial portion of the instances, and in many more cases, we can at least obtain an
upper bound. This is a significant improvement over state of the art maximum common subgraph
solvers, which cannot even fit many of these instances in RAM. Finally, we show that iteratively
increasing $k$ gives a competitive algorithm for the maximum common induced subgraph problem.

\begin{figure}
    \begin{center}
        \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm,node distance=10mm,fill=black}]
            \node[fill=uofgcobalt] (a) {};
            \node (a1) [below left of = a] {};
            \node (b1) [below right of = a] {};
            \node (aa) [below right of = a1] {};

            \node (a2) [below left of = aa,node distance=13mm,xshift=2mm] {};
            \node (aa2) [above left of = a2,node distance=13mm,xshift=-2mm] {};
            \node (ab2) [below left of = a2,node distance=13mm,xshift=-2mm] {};
            \node (ac2) [fill=gray!50,above left of = ab2,node distance=13mm,xshift=-2mm] {};
            \node (b2) [below right of = aa,node distance=13mm,xshift=-2mm] {};
            \node (ba2) [above right of = b2,node distance=13mm,xshift=2mm] {};
            \node (bb2) [below right of = b2,node distance=13mm,xshift=2mm] {};
            \node (bc2) [fill=gray!50,above right of = bb2,node distance=13mm,xshift=2mm] {};

            \node (aaa) [below of =aa, node distance=20mm] {};

            \path[dotted] (a) edge (a1) edge (b1);
            \path (aa) edge (a1) edge (b1) edge (a2) edge (b2)
                (a1) edge (b1)
                (aaa) edge (a2) edge (b2)
                (a2) edge (aa2) edge (ab2)
                (b2) edge (ba2) edge (bb2);
            \path[color=gray!50] (ac2) edge (ab2) edge (aa2)
                (bc2) edge (bb2) edge (ba2);

            \node [cross] {};
        \end{tikzpicture}
        \caption{Example of  ``as much of the pattern'' as possible.}
        \label{fig:subgraphexample}
    \end{center}
\end{figure}
\subsection{Definitions and Notations}

A non-induced subgraph isomorphism from a graph $P$ to a graph $T$ is an injective mapping $P
\rightarrowtail T $ which maps adjacent vertices to adjacent vertices. An induced subgraph
isomorphism $P \hookrightarrow T$ additionally maps non-adjacent vertices to non-adjacent vertices.
We define a $k$-less-subgraph isomorphism from $P$ to $T$ to be a subgraph isomorphism from all but
$k$ vertices of $P$ to $T$; this may be non-induced or induced, written $P \lessnonind{k} T$ and $P
\lessind{k} T$ respectively.

The \emph{complement} of a (simple) graph $P$ is the graph $\shortoverline{P}~$.  Adjacent distinct
vertices in $P$ are non-adjacent in $\shortoverline{P}$ while non-adjacent vertices in $P$ are
adjacent in $\shortoverline{P}$.  The \emph{loop complement} of a graph $P$ is the graph
$\loopcomp{P}$.  The construction of $\loopcomp{P}$ is similar to that of $\bar{P}~$, however
whenever a vertex has a loop in $P$ it does not in $\loopcomp{P}$ and vice versa.  When solving an
induced subgraph isomorphism problem we make use of the loop complement; we care about mapping
vertices with loops to vertices with loops, and vertices without loops to vertices without loops.
The following proposition follows directly from the definitions.

\begin{proposition}\label{prop:comp}
Let $i$ be an assignment of $T$ vertices to $P$ vertices.  Then $i$ satisfies the definition of $P
\hookrightarrow T$ if and only if $i$ satisfies both the definition of $P \rightarrowtail T$ and
of $\loopcomp{P} \rightarrowtail \loopcomp{T}$.
Similarly, $i$ satisfies the definition of $P
\lessind{k} T$ if and only if $i$ satisfies both the definition of $P \lessnonind{k} T$ and of
$\loopcomp{P}
\lessnonind{k} \loopcomp{T}$.
\end{proposition}

A common induced subgraph of graphs $G$ and $H$ is a graph $P$, together with two induced subgraph
isomorphisms to $G$ and $H$; a maximum common induced subgraph is a common induced subgraph with as
many vertices as possible. An induced $k$-less subgraph isomorphism $P \lessind{k} T$ is equivalent
to a common induced subgraph of $P$ and $T$ with $\left|\V(P)\right| - k$ vertices.

If defined similarly, a maximum common non-induced subgraph would allow us to select every vertex in
the smaller of the two graphs, and none of the edges. It is therefore traditional to change the
objective to maximise the number of edges selected, rather than vertices, when a non-induced common
subgraph is sought---this problem is usually called the maximum common \emph{partial} subgraph
problem instead. However, this is not what we will be discussing in this paper: maximum common
subgraph problems are symmetric in their inputs, but when discussing the non-induced case we are
allowing extra edges only in the target graph, not in the pattern.

\begin{example}
Take a pattern graph $P$ with vertices $V=\{p, a, b, c, d\}$ and edges $E=\{(a,d), (a,p), (b,c),
(b,d), (b,p), (c,d), (c,p)\}$, which is shown in \cref{fig:klessexample} along with two target
graphs $T_{1}$ and $T_{2}$. The degree of the vertex $p$ in $P$ is $\deg(p)=3$, the neighbours of
$p$ are $\N(p)=\{a,b,c\}$, the neighbourhood degree sequence of $p$ is $\nds(p)=(3,3,2)$, as the
degrees of the neighbours of $p$ are $\deg(a)=2,\ \deg(b)=3$ and $deg(c)=3$.

We can find a subgraph in $T_{1}$ which is non-induced isomorphic to $P$ (indicated through the bold edges and black vertices in $T_{1}$). The vertex $p$ can be  mapped to $t_{1}$ in the target graph $T_{1}$, where $\deg(t_{1})=5$ and $\nds(t_{1})=(5,5,4,2,2)$.

There is a non-induced mapping of the pattern graph $P$ into $T_{2}$ by removing one vertex of $P$. This removal changes the neighbourhood degree sequence of $p$ in the $k$-less version of $P$, to $\nds(p)=(2,2)$.

\begin{figure}
    \centering
    \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
        \node[fill=black] (p) {};
        \node[fill=black] (b) [above of = p,node distance=10mm] {};
        \node[fill=black] (a) [left of = b] {};
        \node[fill=black] (c) [right of = b] {};
        \node[fill=black] (d) [above of = b,node distance=10mm] {};

        \node[draw=none,right of = p,node distance=3mm,yshift=-1mm] {$p$};
        \node[draw=none,left of = a,node distance=3mm,yshift=-1mm] {$a$};
        \node[draw=none,left of = b,node distance=3mm,yshift=-1mm] {$b$};
        \node[draw=none,right of = c,node distance=3mm,yshift= 1mm] {$c$};
        \node[draw=none,above of = d,node distance=3mm,yshift=1mm] {$d$};

        \node[draw=none, below of = p, node distance=6mm] {$P$};

        \path (p) edge (a) edge (b) edge (c)
            (a) edge (d)
            (b) edge (c) edge (d)
            (c) edge (d);
        \end{tikzpicture}\hspace{0.5cm}\begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
        \node[fill=gray!50] (a1) {};
        \node[fill=black] (b1) [right of = a1] {};
        \node[fill=gray!50] (c1) [right of = b1] {};
        \node[fill=black] (a2) [below of = a1] {};
        \node[fill=black] (b2) [right of = a2] {};
        \node[fill=black] (c2) [right of = b2] {};
        \node[fill=gray!50] (a3) [below of = a2] {};
        \node[fill=black] (b3) [right of = a3] {};
        \node[fill=gray!50] (c3) [right of = b3] {};

        \node[draw=none,right of = b3,node distance=1mm, yshift=-2mm] {$t_{1}$};

        \node[draw=none, below of = b3, node distance=6mm] {$T_{1}$};

        \path[color=gray!50] (a1) edge (b1) edge (a2)
            (b1) edge (c1)
            (c1) edge (c2)
            (a2) edge (b2) edge (a3)
            (c2) edge (c3)
            (a3) edge (b3)
            (b3) edge (c3);
        \path[line width=0.4mm] (b1) edge (a2) edge (b2) edge (c2)
            (a2) edge (b3)
            (b2) edge (c2) edge (b3)
            (c2) edge (b3);
    \end{tikzpicture}\hspace{0.5cm}\begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
        \node[fill=black] (p) {};
        \node[draw=none] (b) [above of = p,node distance=10mm] {};
        \node[fill=black] (a) [left of = b] {};
        \node[fill=black] (c) [right of = b] {};
        \node[fill=black] (d) [above of = b,node distance=10mm] {};

        \node[draw=none, below of = p, node distance=6mm] {$T_{2}$};

        \path (p) edge (a) edge (c)
            (a) edge (d)
            (c) edge (d);
    \end{tikzpicture}
    \caption{$P\rightarrowtail T_{1}$, $P\lessnonind{1} T_{2}$}
    \label{fig:klessexample}
\end{figure}
\end{example}

\subsection{Constraint Models and Algorithms}

For both subgraph isomorphism and maximum common subgraph, constraint programming is the best known
approach\footnote{Note that toolkits are not used in state of the art implementations, to facilitate
better memory layouts and propagation queues, although the algorithms are heavily inspired by
constraint programming.}. For both problems, we create a variable for each vertex in the
pattern graph (the smaller graph, in the case of maximum common subgraph), with domains ranging over
the vertices in the target graph (the larger graph). For maximum common subgraph, each domain is
given an additional $\bot$ value, meaning ``unmapped''.

Constraints are used to ensure that adjacency pairs of vertices are mapped to adjacenct pairs of
vertices. For the maximum common subgraph case, any pair of assignments with $\bot$ as either value
is also permitted.  To handle the induced case, rather than the more usual approach of directly
having constraints for non-adjacent pairs of vertices, we will make use of \cref{prop:comp} by
seeking an $i$ which simultaneously satisfies $P \rightarrowtail T$ and $\loopcomp{P}
\rightarrowtail \loopcomp{T}$. Finally, an all different except $\bot$ constraint ensures
injectivity.

For subgraph isomorphism, this model can be enhanced with domain filtering at the top of search, to
reduce the initial sizes of domains---we discuss this in \cref{section:degreefiltering}. Subgraph
isomorphism also allows us to generate additional implied constraints, which we discuss in
\cref{section:pathfiltering}. Neither of these techniques are valid for the maximum common subgraph
problem.

?? Also the clique encoding.

\subsection{Algorithms for the $k$-Less Problem}

A non-induced $k$-less-subgraph isomorphism from $P$ to $T$ is equivalent to a subgraph isomorphism
between $P$ and $T$, with $k$ extra universally-adjacent vertices added to $T$. However, using such
an approach is not ideal, because it would introduce symmetries; for induced $k$-less-subgraph
isomorphisms, an approach based around adding vertices to $T$ cannot work at all.
Another algorithmic approach could be to try each way of removing $k$ vertices from the pattern
graph, and solving the subgraph isomorphism problem in turn. This approach might be feasible for $k
= 1$, although it would involve a lot of duplication of search effort, but for larger values of $k$
the number of searches which would have to be made would grow very quickly.
Finally, for the induced case we could try adapting maximum common subgraph algorithms to solve the
decision problem. However, both the maximum common subgraph algorithm and the clique encoding
require $O(\left|\V(P)\right|^2\left|\V(T)\right|^2)$ memory, which is extremely problematic on the
instances we wish to work with.

Instead, we introduce a new algorithm, inspired by a state of the art subgraph isomorphism
algorithm. The approach we discuss requires only $O(\left|\V(P)\right|^2\left|\V(T)\right|)$ space
(this can be thousands of times less than maximum common subgraph approaches, once the constant
factors and orders of $T$ in our instances are considered). We list the algorithm in
\cref{algorithm:sip}, explain it fully in \cref{section:algorithm}, and evaluate it in
\cref{section:evaluation}, but first we will discuss our experimental setup, and then detail some
experiments to verify that we can gain effective domain filtering (\cref{section:degreefiltering})
and implied constraints (\cref{section:pathfiltering}) by using subgraph isomorphism style reasoning
about degrees and paths in the $k$-less setting.

\subsection{Experimental Setup} \label{section:setup}

We perform our experiments on systems with dual Intel Xeon CPU E5-2640 processors with 64GBytes of
RAM, running Ubuntu 14.04. Our algorithm was implemented in C++\footnote{URL removed for anonymous
review}, and source code was compiled using
GCC 5.3.0. We used a timeout of 1,000 seconds for each instance. For comparison, we used the
implementations of \citet{DBLP:conf/cp/McCreeshNPS16}.

For datasets, we use the same 5,725 instances used in a recent work on portfolios of algorithms for
subgraph isomorphism \cite{thelionpaper}. This collection includes randomly generated scale-free
graphs \cite{constraints10}, an assortment of real-world graphs of varying sizes \cite{LV02},
randomly generated graph pairs (using bounded degree, regular mesh, and uniform models; all are
satisfiable), segmented images~\cite{pr15,cviu11}, meshes from modelling 3D objects \cite{cviu11},
and graphs close to the satisfiable-unsatisfiable phase transition
\cite{DBLP:conf/ijcai/McCreeshPT16}; all these instances are available in a simple text
format\footnote{\texttt{http://liris.cnrs.fr/csolnon/SIP.html}}.  Note that many of these instances
are \emph{much} larger than were used in the recent comparison of maximum common subgraph algorithms
by \citet{DBLP:conf/cp/McCreeshNPS16}: for unlabelled graphs, they used pairs of graphs with up to
35 vertices in each, whilst our dataset contains graphs with up to 6,671 vertices. For the CP
algorithm, 1,560 of the instances cannot fit in the amount of RAM we have available---we treat these
instances as having timed out. The situation for the clique encoding is even worse, and 3,653 of
these instances do not fit in RAM. In contrast, for our algorithm, every instance fits comfortably
(at least on modern server-class hardware).

\section{Domain Filtering using Degrees}\label{section:degreefiltering}

Let $p$ be a vertex in a graph $P$. The \emph{degree} of $p$, $\deg(p)$, is the number of vertices
to which it is adjacent. The degree of vertices gives us an invariant, which may be used to
eliminate some infeasible values from domains.

\begin{proposition}
    \label{prop:deg}
    Let $p$ be a vertex in $P$ and $t$ a vertex in $T$. For both non-induced and induced
    $k$-less-subgraph isomorphisms, if $p \lessmap{k} t$ then
    $\deg(p) - k \le \deg(t)$.
\end{proposition}

A vertex $q$ which is adjacent to $p$ is called its \emph{neighbour}.  Let $\N(p) = \{ q \in \V(P) :
 p \text{ and } q \text{ are adjacent }\}$ be the set of neighbours of $p$.  The \emph{neighbourhood
degree sequence} (NDS) of $p$, $\nds(p)$, is the (non-ascending) sequence of degrees of its
neighbours.

Let $S = ( s_1 , \ldots , s_n )$ and $T = ( t_1 , \ldots , t_m)$ be two sequences.  We say that $S
\preceq T$ if $n \leq m$ and $\forall s_i \in S$ there exists a distinct $t_j \in T$ with $s_i \leq
t_j$.  When considering a $k$-less-subgraph isomorphism we say that $S \lesspreceq{k}
T$ if $n - k \leq m$ and $\forall s_i \in S \setminus S_k$, where $S_k$ is a subsequence of $S$
containing up to $k$ members of $S$, there exists a distinct $t_j \in T$ with $s_i - k \leq t_j$.

\begin{proposition}\label{prop:nds}
If $p \lessmap{k} t$, then $\nds(p) \lesspreceq{k} \nds(t)$.
\end{proposition}

\Cref{fig:klessexample} illustrates the three cases possible when filtering by neighbourhood degree
sequence.  Removing the top node causes each entry in $\nds(p)$ to be reduced; removing the
left-middle node removes an entry from $\nds(p)$; and removing either of the connected nodes in the
middle layer causes both the size of $\nds(p)$ to be reduced and an entry in $\nds(p)$ to be reduced.

\begin{corollary}
Since both $\nds(p)$ and $\nds(t)$ are non-ascending, without loss of generality we can replace
$S \setminus S_k$ from the definition of $S \lesspreceq{k} T$ with $S[k, n]$.
\end{corollary}

% \begin{figure}
%     \begin{center}
%         \begin{tikzpicture}[every node/.style={circle, inner sep=1mm}]
%             \node[fill=black] (p) {};
%             \node[draw=none,right of = p,node distance=3mm,yshift=-1mm] {$p$};
%             \node[fill=uofglawn] (n2) [above of = p,node distance=12mm] {};
%             \node[fill=uofgpumpkin] (n1) [left of = n2] {};
%             \node[fill=uofglawn] (n3) [right of = n2] {};
%             \node[fill=uofgcobalt] (q) [above of = n2,node distance=12mm] {};
%             \path (p) edge (n1) edge (n2) edge (n3)
%                 (n1) edge (q)
%                 (n2) edge (n3) edge (q)
%                 (n3) edge (q);
%         \end{tikzpicture}
%         \caption{Example of neighbourhood degree filtering cases}
%         \label{fig:nds}
%     \end{center}
% \end{figure}

?? Talk about iterating, what to do when a vertex disappears from all target graphs.

In \cref{figure:ids} we show that this is effective in practice: we show the amount of domain reduction
we can achieve at the top of search by using invariants and a fixed $k$, compared to the search
space size for the maximum common induced subgraph problem. (The results on the non-induced version
show a similar but slightly weaker trend.)

Some other invariants do not translate. For example, another rule which can be effective on regular
graphs involves count the number of neighbours of a vertex which are present in a triangle
\cite{mckay2014practical}. Removing a single vertex can alter this count arbitrarily, so we cannot
make use of this fact.

\begin{figure}
    \centering
    \includegraphics*{gen-graph-ids.pdf}
    \caption{The amount of domain reduction achieved for the induced problem, with increasing values
    of $k$. The $y$ value shows for how many instances we may reduce the initial search space to be
    at most the $x$ proportion of the size the search space would be for the maximum common subgraph
    problem.}\label{figure:ids}
\end{figure}

\section{Filtering During Search Using Paths}\label{section:pathfiltering}

\begin{figure*}[tb]
    \centering
    \includegraphics*{gen-graph-constraints-induced-0.pdf}
    \includegraphics*{gen-graph-constraints-induced-1.pdf}
    \includegraphics*{gen-graph-constraints-induced-2.pdf}
    \includegraphics*{gen-graph-constraints-induced-3.pdf}
    \includegraphics*{gen-graph-constraints-induced-4.pdf}
    \includegraphics*{gen-graph-constraints-induced-5.pdf}

    \caption{For the induced problem, the proportion of pairs of assignments from the filtered
    domains which are not permitted simultaneously, without path constraints on the $x$-axis and
    with path constraints on the $y$-axis, for increasing values of $k$.}\label{figure:constraints}
\end{figure*}

As well as reasoning about degrees, we can also reason about paths.  A \emph{path} between two
vertices $p$ and $q$ is a sequence of edges which can be traversed from $p$ to reach $q$. We define
$\paths(p,q,n)$ to be the number of paths of length $n$ between the vertices $p$ and $q$.

\begin{proposition}\label{prop:paths}
    Let $p,q \in \V(P)$ and $t,u\in \V(T)$. If $p \lessmap{k} t$ and $q \lessmap{k} u$ then
     $\paths(p, q, 2) - k \le \paths(t, u, 2)$.
\end{proposition}

\begin{corollary}
    For a graph $G$, let $G^{n, \ell}$ be the graph with vertex set $\V(G)$. The
    vertices $p$ and $q$ in $G^{n, \ell}$ are adjacent, if there are at least $n$ simple paths of
    length exactly $\ell$ between $p$ and $q$ in $G$. Then any $k$-less-subgraph isomorphism
    $P \lessnonind{k} T$ induces a new $k$-less-subgraph isomorphism
    $P^{n+k, 2}\lessnonind{k} T^{n, 2}$.
\end{corollary}

Unlike in conventional subgraph isomorphism, we cannot extend this filtering to look at paths of
length three: removing a single vertex can delete arbitrarily many such paths. We \emph{could}
instead count paths of any length which are vertex disjoint, although calculating this appears to be
too expensive to be beneficial in practice.

To allow for fast propagation, rather than calculating paths dynamically like
\citet{DBLP:conf/cp/AudemardLMGP14}, we follow the approach of \citet{DBLP:conf/cp/McCreeshP15} and
construct \emph{supplemental graphs}, and find a mapping which is simultaneously a non-induced subgraph
isomorphism between every supplemental graph pair. We use paths of length 2, looking at counts of 1,
2, and 3 in the target graph (and so counts of $1 + k$ up to $3 + k$ in the pattern).  We then
investigate whether this leads to new constraints being generated.

By an assignment, we mean considering mapping a pattern vertex $p$ to a target vertex $t$ (and not
$\bot$) which does not violate any loop constraints. An assignment pair is two assignments with
distinct $p$ and distinct $t$, which we say is permitted if it does not violate any adjacency
constraint. We define the permitted assignment pair ratio to be the proportion of assignment pairs
which are permitted.  Finally, we scatter plot the permitted assignment pair ratio with and without
supplemental graphs\footnote{Because of the large sizes of the domains, we randomly sample one
million pairs rather than considering every pair. In some cases, we have nearly a thousand domains,
each with nearly ten thousand values---a complete quadratic calculation involving even a trivial
arithmetic operation on this would take many hours.}.

For $k = 0$, we see many points above the $x-y$ diagonal, which shows that for many instances, we
are able to create a substantial number of new constraints at the top of search; on the other hand,
there are also points on the diagonal, which shows that sometimes this technique provides no
benefit. (Occasionally, points fall below the $x-y$ diagonal.  This is because the use of
neighbourhood degree sequence reasoning on supplemental graphs can also lead to increased domain
filtering, which could in turn eliminate a higher proportion of forbidden than permitted assignment
pairs.) For $k = 1$ and $k = 2$, the proportion of points above the diagonal diminishes, but we are
still able to create new constraints for many instances. By the time $k = 3$, most of the benefit is
disappearing---although sometimes we are still able to make a difference, and bear in mind that
sometimes adding just one new constrained pair can vastly reduce the search space.

\section{A New Algorithm}\label{section:algorithm}

\begin{algorithm*}[tb]
\DontPrintSemicolon
\begin{multicols}{2}
\nl $\FuncSty{klessSubgraphIsomorphism}$ (\\\hspace*{2em}Graph $\mathcal{P}$, Graph $\mathcal{T}$, Int $k$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \lIf{$\left|\V(\mathcal{P})\right| + k > \left|\V(\mathcal{T})\right|$}{$\KwSty{return}~\KwSty{false}$\label{line:enough}}
    \nl $L \gets \big[ (\mathcal{P},\ \mathcal{T}), (\loopcomp{\mathcal{P}},
        \loopcomp{\mathcal{T}}) \textnormal{~only if we want induced},$
        \\\hspace*{1.0em}$(\mathcal{P}^{1 + k,2},\ \mathcal{T}^{1,2}), (\mathcal{P}^{2 + k,2},\
    \mathcal{T}^{2,2}), (\mathcal{P}^{3 + k,2},\ \mathcal{T}^{3,2}) \big]$ \;
    \nl \ForEach{$v \in \V(\mathcal{P})$}{
        \nl $D_v \gets \V(\mathcal{T})$\;
        \nl \ForEach{$(P,\,T) \in L$}{
            \nl $D_v \gets \{ w \in D_v : $ \\
            \hspace*{4em}$v \sim_P v \Rightarrow w \sim_T w~\wedge$ \\
            \hspace*{4em}$\nds_P(v) \lesspreceq{k} \nds_T(w)\}$ \;
        }
        \nl $D_v \gets D_v \cup k~\textnormal{distinct wildcard values}$
    }
    \nl \lIf{\FuncSty{propagate}(D)}{$\KwSty{return}~\FuncSty{search}(L, D, k)$}
    \nl \lElse{$\KwSty{return}~\KwSty{false}$}
}
\bigskip
\nl $\FuncSty{search}$ (GraphPairs $L$, Domains $D$, Int $k$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \lIf{$D = \emptyset$}{$\KwSty{return}~\KwSty{true}$}
    \nl $D_v \gets \textnormal{the smallest domain in}~D$ \;
    \nl \ForEach {$v' \in D_v~\textnormal{ordered by static degree in}~\mathcal{T}$\label{line:parforeach}}{
        \nl \If{$v'$ \textnormal{is not the first wildcard we have tried}}{
            \nl \KwSty{continue} \;
        }
        \nl $D' \gets \FuncSty{clone}(D)$ \;
        \nl $D'_v \gets \{ v' \} $ \;
        \nl \If{\FuncSty{propagate}(D')}{
            \nl \If{$\FuncSty{search}(L, \{ D \in D' : \left|D\right| > 1 \}, k)$}{
                \nl $\KwSty{return}~\KwSty{true}$ \;
            }
        }
    }
}
\bigskip
\nl $\FuncSty{propagate}$ (GraphPairs $L$, Domains $D$) $\rightarrow$ Bool \;
\nl \Begin{
    \nl \While{\KwSty{true}}{
        \nl \If{\textnormal{no effectively-unit domains (treating all wildcard~values as a single
        value) remain}}{
            \nl \If{$\KwSty{not}~\FuncSty{allDifferent}(D)$}{
                \nl $\KwSty{return}~\KwSty{false}$ \;
            }
        }
        \nl \If{\textnormal{no effectively-unit domains remain}}{
            \nl $\KwSty{return}~\KwSty{true}$ \;
        }
        \nl $D_v \gets \textnormal{an effectively-unit domain from}~D$ \;
        \nl $v' \gets $ the single value in $D_v$, or an arbitrary wildcard~value if there is more
        than one \;
        \nl \ForEach{$D_w \in D \setminus \{ D_v \}$}{
            \nl $D_w \gets D_w \setminus \{ v' \}$ \label{line:removev} \;
            \nl \ForEach{$(P,\,T) \in L$}{
                \nl \If{$v \sim_P w$}{
                    \nl $D_w \gets D_w \cap \left(\N_T(v') \cup \textnormal{wildcards}\right)$
                }
            }
            \nl \lIf{$D_w = \emptyset$}{$\KwSty{return}~\KwSty{false}$}
        }
    }
}
\bigskip
\nl $\FuncSty{allDifferent}$ (Domains $D$) $\rightarrow$ Bool \;
\nl \Begin{
\nl $(H,\,A,\,n) \gets (\emptyset,\,\emptyset,\,\emptyset,\,0)$ \;
\nl \ForEach{$D_v \in D$ \textnormal{from smallest to largest\label{line:eachdomain}}}{
    \nl $D_v \gets D_v \setminus H$ \label{line:elimhall} \;
    \nl $(A,\,n) \gets (A \cup D_v,\,n + 1)$ \label{line:acc} \;
    \nl \lIf{$D_v = \emptyset~\KwSty{or}~\left|A\right| < n$}{$\KwSty{return}~\KwSty{false}$\label{line:failhall}}
    \nl \lIf{$\left|A\right| = n$}{$(H,\,A,\,n) \gets (H \cup A,\,\emptyset,\,0)$\label{line:hall}}
}
\nl $\KwSty{return}~\KwSty{true}$ \;
}
\end{multicols}
\vspace*{0.4cm}% bug in algorithm2e + multicol: incorrect space left above the vertical line
\caption{An algorithm for the $k$-less subgraph isomorphism problem.}
\label{algorithm:sip}
\end{algorithm*}

\cref{algorithm:sip}. This is derived from a subgraph isomorphism due to
\citet{DBLP:conf/cp/McCreeshP15}, which in a recent evaluation of algorithm portfolios
\citep{thelionpaper} was the single strongest solver. As well as the modifications to solve $k$-less
and induced problems, we have made two simplifications: we do not use conflict-directed backjumping,
and we do not use iterated label filtering to recalculate neighbourhood degree sequences if vertices
are not present in any target domain after construction---in preliminary experiments, these
techniques were not beneficial for $k > 0$.

?? Discuss how we did wildcards, effectively unit domains. The all diff propagator.

?? Discuss how we do induced. We do not look at paths in the loop complements, since this is
expensive to calculate and provides little benefit in practice on the instances we considered.

\section{Empirical Evaluation}\label{section:evaluation}

\begin{figure*}[tb]
    \centering
    \includegraphics*{gen-graph-runtimes.pdf}
    \caption{In the first two plots, the cumulative number of instances solved over time, for the
        induced and non-induced problems, with different values of $k$. We also show the results of
        iteratively decreasing $k$ until a solution is found, and in the induced case, the
        performance of two leading maximum common subgraph algorithms. In the third plot, results
        comparing our algorithm with iteratively decreasing $k$ to other approaches on maximum
        common induced subgraph instances.}\label{figure:runtimes}
\end{figure*}

\begin{figure*}[tb]
    \centering
    \includegraphics*{gen-graph-which-k-by-family.pdf}
    \caption{The proportion of instances, in different families, which become satisfiable for
    increasing values of $k$. The larger instances are those where we can prove unsatisfiability for
    $k = 5$, whilst the gap between the top of the bar and the top of the graph is the fraction of
    instances where a timeout was reached for at least one value of $k$.}\label{figure:which-k}
\end{figure*}

We now evaluate our algorithm and show that it is effective in practice, even on the larger subgraph
isomorphism instances. In \cref{figure:runtimes} we give cumulative distribution plots for the
induced and non-induced problems, with $k$ ranging from $0$ to $5$ (we discuss the dotted lines in
this plot below). The results are strong: with $k = 0$ we may solve nearly every instance, whilst
even at $k = 5$ we can solve over $4,000$ instances in both variants.

But are we learning anything about the results---that is, are there instances for which $k = 0$ is
unsatisfiable, but that are satisfiable for small $k$? For the problem families which do not consist
entirely of satisfiable instances, we plot this in \cref{figure:which-k}. In the ``phase'' family,
which consists of instances crafted to be extremely difficult to solve, we are not able to answer
this question, and in the ``scalefree'' family we see no satisfiable instances with low but not zero
$k$ (this is simply due to there being many vertices with loops in some pattern graphs, but no loops
at all in the targets). However, in several of the remaining families we can over double the number
of instances for which exact solutions are known, and gain upper bounds on many more. This is
particularly interesting for the ``images-CVIU11'' family, where the size of the solution has a
direct real-world interpretation in terms of closeness of image matching.

\subsection{Solving from the Top Down}

What would happen if we used our approach to solve the maximum common induced problem? We could
simply start at $k = 0$, and increase $k$ until a solution is found. This would be tackling the
problem in the opposite direction from existing approaches, which work by attempting to construct
larger and larger solutions. The dotted lines in \cref{figure:runtimes} show this approach: the $k
\downarrow$ is our algorithm, whilst ``FC'' and ``clique'' are the two algorithms discussed in ??.
Bearing in mind the disclaimer in \cref{section:setup} regarding instances not fitting in RAM, we
see that our approach is able to close over twice as many of these instances as the previous state
of the art. (The same conclusion holds even if every instance which is satisfiable with $k = 0$ is
removed from the dataset.)

What about if we use instances designed for the maximum common subgraph problem? In the third plot
in \cref{figure:runtimes} we again compare these approaches, but using the first ten instances from
each family of a standard benchmark set
\citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}, selecting undirected, unlabelled
graphs with up to 50 vertices---this gives us a total of 4,110 instances. In these instances the
number of vertices in the pattern and target graphs is the same, which is not ideal for our
algorithm (although our invariants are still effective in many cases). Nonetheless, we are by a
small margin the single strongest solver. Interestingly, our algorithm tends to have complementary
performance to the other two approaches (which are much more correlated), which suggests that there
is scope for an algorithm which runs both an upper bound and a constructive, lower bound approach
simultaneously or in parallel, stopping when the two bounds meet in the middle.

\section{Conclusion}

We looked at . We saw that we could modify invariants, and that doing so still provided additional
domain reduction and filtering power.

Relaxed form of the problem which is easier to solve. Far from making maximum common subgraph scale
as well as subgraph isomorphism, but we can at least get somewhere.

Future work.

\section*{Acknowledgements}

We'd like to thank some people whose names we've removed for anonymous review for providing us with
source code and problem instances, and someone at one of our institutions for inspiration.

\bibliographystyle{named}
\bibliography{paper}

\clearpage
\appendix

\section{Proofs Omitted From Main Text}

\begin{proof}[Proof of \cref{prop:deg}]
Let $p$ be a vertex in $P$, and $t$ a vertex in $T$, with $p\mapsto t$. Then by the definition of subgraph isomorphisms, $\deg(p) \le \deg(t)$. Let $P'$ be $P$ less $k$ vertices and $p$ a vertex in $P'$. Then
\[
\deg_{P}(p)-k \le \deg_{P'}(p) \le \deg_{P}(p) \le \deg(t). \qedhere
\]
\end{proof}

\begin{proof}[Proof of \cref{prop:nds}]
Let $p \lessmap{k} t$.  Then $\deg(p) - k \leq \deg(q)$, by \cref{prop:deg}, which
implies that $\left|\nds(p)\right| -k \leq \left| \nds(t) \right| $.

Also, $p \lessmap{k} t$ implies that for each $q \in N(p) \setminus P_k$, where $P_k$
is some subset of $P$ with $\left| P_k \right| \leq k$, $q \lessmap{k} u$, where $u \in N(t)$ and
each $u$ is distinct.  Then $\deg(q) - k \leq \deg(u)$, by \cref{prop:deg}.  Hence $\nds(p)
\lesspreceq{k} \nds(t)$.
\end{proof}

\begin{proof}[Proof of \cref{prop:paths}]
Let $P_{2}$ be the set of all paths of length two between $p$ and $q$,
\begin{align*}
P_{2} = & \{((p,x),(x,q)) : (p,x),(x,q)\in E(P) \} \\
 = & \{((p,x),(x,q)) : x\in \N(p) \text{ and } x\in \N(q) \} \\
 = & \{((p,x),(x,q)) : x\in \N(p)\cap \N(q) \} .
\end{align*}
So $\paths(p,q,2) = \left| \N(p)\cap \N(q) \right|  \leq \deg(p)$.

If we remove up to $k$ vertices from the neighbourhood of $p$, it will impact the intersection of the neighbourhoods of $p$ and $q$, and by Proposition~\ref{prop:deg}, as $p\lessmap{k}t$
\[
\left| \N(p)\cap \N(q)\right| - k \leq \deg(p) - k \leq \deg(t).
\]
As $\left|\N(p)\cap \N(q)\right|\leq \deg(p) \leq \left|\N(t)\cap \N(u)\right|\leq \deg(t)$, we have
\[
\left|\N(p)\cap \N(q)\right| - k \leq \deg(p) - k\leq \left|\N(t)\cap \N(u)\right|\leq \deg(t)
\]
which is
\[
\paths(p, q, 2) - k \leq \paths(t, u, 2). \qedhere
\]
\end{proof}

\end{document}