### 2017-01-31 slide/slide.tex

\documentclass{beamer}
\usepackage[export]{adjustbox}

\usepackage{tikz}
\usetikzlibrary{automata,shapes,arrows,positioning,fit}

\input{talkpreamble}

\title[Between Subgraph Isomorphism and Maximum Common Subgraph]{Between Subgraph Isomorphism and \\ Maximum Common Subgraph}
\author[Ruth Hoffmann, Ciaran McCreesh and Craig Reilly]{Ruth Hoffmann, Ciaran McCreesh and  \textbf{Craig Reilly}}

\begin{document}

    \begin{frame}[plain,noframenumbering]
        \titlepage
        \vspace{2.2cm}
        \begin{figure}[b]
            \centering
            \begin{tikzpicture}[scale=0.58]%[every node/.style={circle, inner sep=0.5mm,node
                %distance=10mm,fill=black}, scale=0.2]

                \begin{scope}[xshift=-1cm, yshift=-3cm, color=black]
                \node at (-2, 1) {$P$:};
                \node [fill, circle, inner sep=1.5pt] (a) at (0,0) {};
                \node [fill, circle, inner sep=1.5pt] (b) at (-1,1) {};
                \node [fill, circle, inner sep=1.5pt] (c) at (0,1) {};
                \node [fill, circle, inner sep=1.5pt] (d) at (1,1) {};
                \node [fill, circle, inner sep=1.5pt] (e) at (0,2) {};

                \path[solid] (a) edge (b) edge (c) edge (d)
                    (e) edge (b) edge (c) edge (d);
                \end{scope}

                \begin{scope}[xshift=4cm, yshift=-3cm, color=black]
                \node at (-2, 1) {$T$:};
                \node [fill, circle, inner sep=1.5pt] (a) at (0,0) {};
                \node [fill, circle, inner sep=1.5pt] (b) at (-1,1) {};
                \node [fill, circle, inner sep=1.5pt] (c) at (1,2) {};
                \node [fill, circle, inner sep=1.5pt] (d) at (1,1) {};
                \node [fill, circle, inner sep=1.5pt] (e) at (0,2) {};
                \node [fill, circle, inner sep=1.5pt] (f) at (1,0) {};
                \node [fill, circle, inner sep=1.5pt] (g) at (2,0) {};
                \node [fill, circle, inner sep=1.5pt] (h) at (2,2) {};

                \path[solid] (a) edge (b) edge (d)
                    (e) edge (b) edge (d)
                    (d) edge (c) edge (f)
                    (c) edge (h)
                    (f) edge (g);
                \end{scope}

            \end{tikzpicture}
        \end{figure}

        \begin{figure}[t]
            \centering
            \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
                \begin{scope}
                \node[fill=black] (p) {};
                \node[fill=black] (b) [above of = p,node distance=5mm] {};
                \node[fill=black] (a) [left of = b, node distance=7mm] {};
                \node[fill=black] (c) [right of = b, node distance=7mm] {};
                \node[fill=black] (d) [above of = b,node distance=5mm] {};

                \node[draw=none,right of = p,node distance=3mm,yshift=-0.5mm] {$p$};
                \node[draw=none,left of = a,node distance=3mm,anchor=base] {$a$};
                \node[draw=none,left of = b,node distance=1.8mm,anchor=base] {$b$};
                \node[draw=none,right of = c,node distance=2mm,anchor=base] {$c$};
                \node[draw=none,right of = d,node distance=3mm,anchor=base] {$d$};

                \node[draw=none, left of = a, node distance=8mm] {$P$:};

                \path (p) edge (a) edge (b) edge (c)
                    (a) edge (d)
                    (b) edge (c) edge (d)
                    (c) edge (d);

                \node[fill=gray!50] (a1) [right of = d, node distance = 2.2cm] {};
                \node[fill=black] (b1) [right of = a1, node distance=7mm] {};
                \node[fill=gray!50] (c1) [right of = b1, node distance=7mm] {};
                \node[fill=black] (a2) [below of = a1, node distance=5mm] {};
                \node[fill=black] (b2) [right of = a2, node distance=7mm] {};
                \node[fill=black] (c2) [right of = b2, node distance=7mm] {};
                \node[fill=gray!50] (a3) [below of = a2, node distance=5mm] {};
                \node[fill=black] (b3) [right of = a3, node distance=7mm] {};
                \node[fill=gray!50] (c3) [right of = b3, node distance=7mm] {};

                \node[draw=none,right of = b3,node distance=1mm, yshift=-2mm] {$q$};

                \node[draw=none, left of = a2, node distance=6mm] {$T$:};

                \path[color=gray!50] (a1) edge (b1) edge (a2)
                    (b1) edge (c1)
                    (c1) edge (c2)
                    (a2) edge (b2) edge (a3)
                    (c2) edge (c3)
                    (a3) edge (b3)
                    (b3) edge (c3);
                \path[line width=0.4mm] (b1) edge (a2) edge (b2) edge (c2)
                    (a2) edge (b3)
                    (b2) edge (c2) edge (b3)
                    (c2) edge (b3);

                \node[draw=none] (b) [right of = c2, node distance=2cm] {};
                \node[fill=black] (p) [below of = b, node distance=5mm] {};
                \node[fill=black] (a) [left of = b, node distance=7mm] {};
                \node[fill=black] (c) [right of = b, node distance=7mm] {};
                \node[fill=black] (d) [above of = b,node distance=5mm] {};

                \node[draw=none, left of = a, node distance=5mm] {$U$:};

                \path (p) edge (a) edge (c)
                    (a) edge (d)
                    (c) edge (d);
                \end{scope}
            \end{tikzpicture}
        \end{figure}
        \begin{figure}[tb]
            \centering
            \begin{tikzpicture}[scale=0.5]%[every node/.style={circle, inner sep=0.5mm,node
                %distance=10mm,fill=black}, scale=0.2]

                \begin{scope}[xshift=-1cm, yshift=-3cm, color=black]
                \node at (-1,0) {$P$:};
                \node [fill, circle, inner sep=1.5pt] (a) at (0,0) {};
                \node [fill, circle, inner sep=1.5pt] (b) at (1,-1) {};
                \node [fill, circle, inner sep=1.5pt] (c) at (1,0) {};
                \node [fill, circle, inner sep=1.5pt] (d) at (1,1) {};
                \node [fill, circle, inner sep=1.5pt] (e) at (2,0) {};
                \node [fill, circle, inner sep=1.5pt] (f) at (3,0) {};
                \node[draw=none,below of = e,node distance=3mm,anchor=base] {$p$};

                \path[solid] (a) edge (b) edge (c) edge (d)
                    (b) edge (e)
                    (c) edge (e)
                    (d) edge (e)
                    (e) edge (f);
                \end{scope}

                \begin{scope}[xshift=6cm, yshift=-3cm, color=black]
                \node at (-1,0) {$T$:};
                \node [fill, circle, inner sep=1.5pt] (a) at (0,0) {};
                \node [fill, circle, inner sep=1.5pt] (b) at (1,-1) {};
                \node [fill, circle, inner sep=1.5pt] (c) at (1,0) {};
                \node [fill, circle, inner sep=1.5pt] (d) at (1,1) {};
                %\node [fill, circle, inner sep=1.5pt] (e) at (2,0) {};
                \node [fill, circle, inner sep=1.5pt] (f) at (3,0) {};

                \path[solid] (a) edge (b) edge (c) edge (d);
                %    (b) edge (e)
                %    (c) edge (e)
                %    (d) edge (e)
                %    (e) edge (f);
                \end{scope}

            \end{tikzpicture}
        \end{figure}
    \end{frame}

\end{document}